Multicast File Transfer Protocol

### 2011-07-13 papers/paper1/paper.tex
another revision of Model

Moving data across a network can take a variety of forms. These forms correspond to different user needs and consequently different protocols. Multicast File Transfer Protocol (MFTP) addresses the need to move immutable data between many different nodes in a fully distributed network. 
%% This is useful for 1) device discovery 2) File transfer in the absence of many technologies (including internet). the only thing that is necessary is computers with routing and broadcast/multicast enabled 3) maximizing the reach of a gossip style network
\paragraph {Components.}
The protocol employs a number of conceptually simple structures--a reflection of its overall simplicity, both in design and in purpose.
One of these structures is called a File. 
Files are a sequence of bytes with a length and a user-defined type; they are broken down into equally sized pieces called fragments.  
Each File has a respective File ID, which is a structure that contains a hash of the file content, length, and type. 
Each File ID is unique to its File.
From these different File and File IDs a matching concept is established: given a File $f_1$ and $f_2$, one can define a matching predicate $\mu$.
Finally, there are three types of Files which are conceptually important to the protocol, the first of which being a request.
A request contains a File ID and a set of contiguous regions.  In our implementation, the set of contiguous regions was a set of fragment indices, but one can envision the request to contain anything that defines a set.
%%A request contains a File ID and a set of fragment indices. 
%%We implemented requests as the File ID and a set of contiguous regions.  
%%However, one can envision the request to contain anything that defines a set.
The second type is a fragment, which consists of a File ID, fragment index, and the data itself.
Last of all is a match, which consists of a File ID and a set of matching File IDs.
\paragraph {Component Interactions.}
In a simple scenario, if one gets a request for a File one has fragments for, one can send those fragments out.
Fragments are periodically broadcasted to indicate what Files exist in the network.
A client can use requests to pull for File fragments, which can span from individual fragments to the whole file.
Two different matching steps are established.
The first uses $\mu_0$ to take two File ID's and determine whether there could be a match between the two Files.  
If this is true, the second uses $\mu_1$ to take two Files and determine whether or not there is a match.
If $\mu_1$($f_1$,$f_2$) is true, $\mu_0$($fid_1$,$fid_2$) must be true.
However, if $\mu_0$($fid_1$,$fid_2$) is true, $\mu_1$($f_1$,$f_2$) does \emph{not} have to be true.

>>>>

% jrw:  Move this to the introduction.
Moving data across a network can take a variety of forms. These forms correspond to different user needs and consequently different protocols. Multicast File Transfer Protocol (MFTP) addresses the need to move immutable data between many different nodes in a fully distributed network. 
%% This is useful for 1) device discovery 2) File transfer in the absence of many technologies (including internet). the only thing that is necessary is computers with routing and broadcast/multicast enabled 3) maximizing the reach of a gossip style network

\paragraph {Files.}
The main concept in MFTP is that of a \emph{file}.
Files are a sequence of bytes with a length and a user-defined type.
They are broken down into equally sized pieces called \emph{fragments}.  
Each file is uniquely identified by its \emph{file ID}, a structure that contains a hash of the file content, the file's length, and the file's type.
Given two files $f_1$ and $f_2$, one can define a matching predicate $\mu$.

\paragraph{Messages.}
File transfer takes place via \emph{messages}.
There are three types of messages in the protocol, the first of which being a \emph{request}.
A \emph{request} contains a file ID and a set of fragment indices.
We implemented requests as the file ID and a set of contiguous regions.  
However, one can envision the request to contain anything that defines a set.
The second message type is a \emph{fragment}, which consists of a file ID, a fragment index, and the data comprising that fragment of the file.
Last is a \emph{match}, which consists of a file ID and a set of matching file IDs.

\paragraph {Tasks.}
MFTP is a peer-to-peer protocol since all participants can act as both a client and a server.
In a simple scenario, if a peer gets a request for a file it has fragments for, it can send those fragments out.
Fragments are periodically broadcast even in the absence of requests to indicate what files exist in the network.
A peer can use requests to poll for fragments of a file which can span from individual fragments to the whole file.
A peer which has successfully downloaded any number of fragments of a file can then serve those fragments.
Matching is a two-step process.
The first step applies a matching function $\mu_0$ to two file IDs to determine whether a match is possible between the two files.  
The second step applies a matching function $\mu_1$ to the two files to determine whether there is a match.
If $\mu_1$($f_1$,$f_2$) is true, $\mu_0$($fid_1$,$fid_2$) must be true.
However, if $\mu_0$($fid_1$,$fid_2$) is true, $\mu_1$($f_1$,$f_2$) does \emph{not} have to be true.
